\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{uniinput}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fullpage}

\newcommand{\imp}{\supset}
\newcommand{\I}{\mathcal{I}}
\newcommand{\Ib}{\I_{¬ F}}
\newcommand{\It}{\I_{F}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\Lf}{\L_F}


\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}

\newtheorem{definition}[theorem]{Definition}

\begin{document}
\section{coNP-completeness of TAUT}

Let $\L$ be a fixed propositional language, i.e. a set of propositional atoms, that does not contain the symbol $F$. Let $\Lf = \L ∪ \{F\}$.
\begin{definition}
 Let $φ$ a propositional formula in the language $\L$. We define a translation $'$ into the positive fragment of propositional logic:
 \begin{enumerate}
  \item For $φ \equiv A$: $φ' \equiv A$.
  \item For $φ \equiv \bot$: $φ' \equiv F$.
  \item For $\circ \in \{∨,∧,\imp\}$ and $φ \equiv ψ_1 \circ ψ_2$: $φ' \equiv ψ_1' \circ ψ_2'$.
  \item For $φ \equiv ¬ψ$: $φ' \equiv ψ' \imp F$.
 \end{enumerate}
\end{definition}

Clearly, the range of $'$ contains no instances of $¬$ or $\bot$.

If $\I$ is an interpretation of $\L$, let $\Ib = \I ∪ \{ F \mapsto \bot\}$ and $\It = \I ∪ \{F \mapsto \top\}$ be interpretations of $\Lf$.
\begin{lemma}
 Let $\I$ be an interpretation of $\L$ and $φ$ an $\L$-formula.
 \begin{enumerate}
  \item If $\I \vDash φ$ then $\Ib \vDash φ'$ and $\It \vDash φ'$.
  \item If $\I \not\vDash φ$ then $\Ib \vDash φ' \imp F$.
 \end{enumerate}
\end{lemma}
\begin{proof}
 By induction.
 \begin{itemize}
  \item $φ \equiv A$: trivial.
  \item $φ \equiv \bot$: trivial.
  \item $φ \equiv ψ_1 ∧ ψ_2$: Assume $\I \vDash φ$. Then $\I \vDash ψ_1$ and $\I \vDash ψ_2$. By induction, $\Ib \vDash ψ_i'$ and $\It \vDash ψ_i'$ for $i \in \{1,2\}$ and hence $\Ib \vDash ψ_1' ∧ ψ_2 \equiv φ' $ and $\It \vDash ψ_1' ∧ ψ_2' \equiv φ'$.
  
  Assume $\I \not \vDash φ$. Then $\I \not \vDash ψ_1$ or $\I \not \vDash ψ_2$. Say $\I \not \vDash ψ_1$ w.l.o.g. Then by induction, $\Ib \vDash ψ_1' \imp F$ and hence $\Ib \vDash (ψ_1' ∧ ψ_2') \imp F \equiv φ' \imp F$.
  \item $φ \equiv ψ_1 ∨ ψ_2, φ \equiv ψ_1 \imp ψ_2$: Analogous to the previous case.
  \item $φ \equiv ¬ ψ$: Assume $\I \vDash φ$, then $\I \not \vDash ψ$. By induction $\Ib \vDash ψ' \imp F \equiv φ'$. $\It \vDash ψ' \imp F \equiv φ'$ is trivially true.
  
  On the other hand, if $\I \not \vDash φ$, then $\I \vDash ψ$. By induction, $\Ib \vDash ψ'$. It follows that $\Ib \vDash (ψ' \imp F) \imp F \equiv φ \imp F$.
 \end{itemize}
\end{proof}
\begin{theorem}
 $φ$ is tautological iff $φ'$ is tautological.
\end{theorem}
\begin{proof}
If $φ$ is tautological, then $\I \vDash φ$ for every $\L$-interpretation $\I$. Since every $\Lf$-interpretation is an $\L$-interpretation extended with either $ F \mapsto \bot$ or $ F \mapsto \top$, we obtain from the lemma that $\I' \vDash φ'$ for every $\Lf$-interpretation $\I'$. Hence $φ'$ is tautological.

Conversely, if $φ$ is not tautological, there is an $\L$-interpretation $\I$ such that $\I \not \vDash φ$. By the lemma, $\Ib \vDash φ' \imp F$ and since $\Ib(F) = \bot$, this means $\Ib \not \vDash φ'$. Hence $φ'$ is not tautological.
 \end{proof}
 
 We have thus shown that for each instance of the tautology problem of propositional logic, we can construct an equivalent instance of the tautology problem for the positive fragment of propositional logic. This construction is obviously possible in polynomial time. In other words, we have reduced the first problem to the second. Since the first problem is coNP-complete, so is the second.

\end{document}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}

\newtheorem{definition}[theorem]{Definition}
