\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{uniinput}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fullpage}
\usepackage{gslist}
\usepackage{gsgraph}

\newcommand{\imp}{\supset}
\newcommand{\I}{\mathcal{I}}
\newcommand{\K}{\textbf{K}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\Ib}{\I_{¬ F}}
\newcommand{\It}{\I_{F}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\Lf}{\L_F}
\newcommand{\Dia}{\Diamond}
\newcommand{\M}{\mathcal{M}}


\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem*{problem*}{Problem}

\theoremstyle{definition}

\newtheorem{definition}[theorem]{Definition}

\title{NCL exercises 6}
\author{Sebastian Zivota}

\begin{document}
\maketitle

\section*{Exercise 46}

\begin{problem*}
Show that the following formulas are BHK-valid:

\begin{enumerate}
  \setcounter{enumi}{2}
 \item $A ∧ B \imp B ∧ A$
\item $A ∨ B \imp B ∨ A$
\item $A \imp (B \imp A)$
\item $\bot \imp A$
\end{enumerate}
\end{problem*}
\begin{enumerate}
\setcounter{enumi}{2}
 \item $π \triangleright \{A ∧ B \imp B ∧ A \}$ means that $π$ is an algorithm $μ$ that transforms a proof $δ \triangleright \{A ∧ B\}$ into a proof $ρ \triangleright \{B ∧ A\}$. $δ \triangleright \{A ∧ B\}$ in turn means $δ = (α,β)$ with $α \triangleright \{A\}, β \triangleright \{B\}$. 
 Obviously, $(β,α) \triangleright B ∧ A$, so we can just let $μ$ be the function that swaps a pair.
 \item We are again looking for a function. Let $μ = \begin{cases}
                                                            (left, α)&\mapsto (right,α),\\
                                                            (right,β)&\mapsto (left,β)
                                                           \end{cases}$.
\item We need an algorithm that, given a proof of $A$, returns another algorithm that transforms any proof of $B$ into a proof of $A$. The lambda term $\mathbf{K}=λx.λy.x$ describes such an algorithm. Expressed in natural language,
given a proof $δ$ of $A$ we can simply transform any proof $ρ$ of $B$ into $δ$.
\item We need an algorithm that can transform any proof of $\bot$ into a proof of $A$. $\mathbf{I} = \lambda x.x$ is such an algorithm: since there are no proofs of $\bot$, the required property is trivially fulfilled.
 \end{enumerate}

\section*{Exercise 47}

\begin{problem*}
Show that the following formulas are BHK-valid:

\begin{enumerate}
  \setcounter{enumi}{6}
 \item $(A \imp (B \imp C)) \imp ((A ∧ B) \imp C)$
\item $¬(A ∨ B) \imp ¬A ∧ ¬B$
\item $(A ∧(B ∨ C)) \imp ((A ∧ B) ∨ (A ∧ C))$
\item $((A ∧ B) ∨ (A ∧ C)) \imp (A ∧(B ∨ C))$
\end{enumerate}
\end{problem*}
\begin{enumerate}
\setcounter{enumi}{6}
 \item We start with a function $δ$ that, for any proof $α \triangleright \{A\}$, produces another algorithm $δ(α)$ that transforms a proof $β \triangleright \{B\}$ into a proof $γ \triangleright \{C\}$. We need an algorithm that
 transforms $(α,β)$ into $γ$. $(α,β) \mapsto δ(α)(β)$ achieves this. In other words, we uncurry the function $δ$.
 \item Suppose we have an algorithm $δ$ that tranforms any proof of $A ∨ B$---that is, any $(left, α)$ with $α \triangleright \{A\}$ or $(right,β)$ with $β \triangleright \{B\}$---into a proof of $\bot$. Let $ρ_1 = λx.δ (left,x)$ and $ρ_2 =
 λx.δ (right, x)$. Then $ρ = (ρ_1,ρ_2)$ is the solution. This works just as well if the negation is replaced with $\imp C$.
 \item The solution is the function $(α,(σ,δ)) \mapsto (σ, (α, δ))$.
\item The solution is $(σ,(α,δ)) \mapsto (α,(σ,δ))$, i.e. the inverse of the previous function.
 \end{enumerate}

 \section*{Exercise 48}
 \begin{problem*}
  Argue informally that the following formulas are intuitionistically invalid:
  \begin{enumerate}
  \item $¬¬A \imp A$
  \item $(A \imp B)  ∨ (B \imp A)$
  \item $¬(A∧B) \imp (¬A ∨ ¬B)$
   \item $¬\forall x A(x) \imp \exists x ¬A(x)$
   \item $((A \imp B) ∧ (¬A \imp B)) \imp B$
  \end{enumerate}
 \end{problem*}
\begin{enumerate}
\item $¬¬A$ intuitively means ``every proof that $A$ is absurd is itself absurd''. This does not give any indication how we would obtain a proof of $A$. An alternative argument: Given double-negation elimination, it is not very hard to derive the law of excluded middle, which we know is not valid.
\item If this formula were valid, then one of its disjuncts would already be valid (and we would know which one), which is obviously not the case.
\item The left-hand side asserts that the combination of $A$ and $B$ is absurd (leads to a contradiction), while the right-hand side asserts that either $A$ or $B$ already leads to a contradiction on its own (and we know which one it is). This is obviously much stronger than the left-hand side.
 \item This the first-order version of the previous example. $¬\forall x A(x)$ says that it is absurd that all elements satisfy $A$. $\exists x ¬A(x)$ asserts that we can in fact construct a concrete element such that it is absurd that that element satisfies $A$, which intuitionistically is stronger.
 \item This formula corresponds to an argument of the form ``if $A$ is true, we can deduce $B$, and if $A$ is false, we can also deduce $B$. Therefore $B$''. This presupposes that $A$ is either true or false, i.e. the law of excluded middle, which we know is false.
\end{enumerate}



\end{document}